Knowledge Representation for Service Robots: Supply and Demand

### 2014-04-16 krs-survey.tex
Summarized the 'features' section in 2 pages

@@

The paper is organized in four main sections.

The next section proposes a prototypical scenario of a robot acting in a rich
semantic environment. We will not survey knowledge representation systems based
on this scenario: we merely use it to ground the selection and illustrate the
set of features that define the dimensions of a knowledge representation system
for robotics.

Section~\ref{sect|features} presents these features in detail, and organize
them into a typology.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A Typology of Knowledge Representation Requirements for Robotics}
\label{sect|features}

We now want to organize requirements from this imaginary cooking scenario where
numerous desirable features for a knowledge representation system for robotics
have been identified without any particular order.

@@

As previously mentioned, we propose six main categories.

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    %[edge from parent fork east]
    %every node/.style={fill=black!30,rounded corners},
    %[parent anchor=east,child anchor=west,grow=east,
    %edge from parent/.style={thick,draw}]
    \node [taxon] {Dimensions}
        child {node [taxon] {{\bf F}. Instantiation}}
        child {node [taxon] {{\bf E}. Integration}}
        child {node [taxon] {{\bf D}. Acquisition}}
        child {node [taxon] {{\bf C}. Reasoning}}
        child {node [taxon] {{\bf B}. Representation}}
        child {node [taxon] {{\bf A}. Expressiveness}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

Table~\ref{table|contribution-by-systems}, at the end of the article,
summarizes all these features with the main domain of contribution of each
surveyed KRS.

@@

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf A. Expressiveness}
            child {node [taxon] {{\bf A.5}. Meta-cognition}}
            child {node [taxon] {{\bf A.4}. Uncertainty}}
            child {node [taxon] {{\bf A.3}. OWA/CWA}}
            child {node [taxon] {{\bf A.2}. Expressive power}}
            child {node [taxon] {{\bf A.1}. Logic formalism}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

@@

\subsubsection{Main logic formalisms}

The main role of a knowledge representation system is to provide an adequate
representation system to formally store facts and concepts that could be informally
described in natural language.

Formal logic aims at providing such a representation system with the added
value of providing a tractable support for inference and reasoning.

Most (but not all) of the systems we have surveyed rely on a particular logic
formalism. The choice of the formalism has a strong impact, on one side, on the
range of ideas that can be expressed conveniently (\emph{practical
expressiveness}) or at all (\emph{theoretical expressiveness}), on the other
side, on the ability to solve the fundamental inference problem (called
\emph{satisfiability}: is a given logical sentence true in my model?) in a
tractable manner.

A large number of logic formalisms do exist and we briefly present below the most
relevant ones for systems actually deployed in robotic architectures.

\emph{Predicate logic} is the family of logic formalisms the most commonly
found in knowledge representation. It distinguishes itself from the simpler
\emph{propositional logic} by the use of quantification to increase generality.
\emph{First-order logic} (FOL) is the subpart of \emph{predicate logic} where the
objects of \emph{predicates} (or \emph{formulae}) are simple \emph{terms},
while in \emph{higher-order logics}, predicates can be themselves objects of
other predicates.

\emph{Horn clauses} are an important subset of FOL because the satisfiability
of a set of such clauses is a $P$-complete problem (\ie practically tractable).
A Horn clause is a disjunction of literals (a \emph{clause}) with at most one
positive literal: $\neg p \lor \neg q \lor \cdots \lor \neg t \lor u$, which
can also be represented as $(p \land q \land \cdots \land t) \rightarrow u$.
Important logic programming languages like Prolog are based on Horn clauses.

The family of \emph{Description Logics}~\cite{Baader2008} also plays an
important role. It is also a subset of the first-order logic, with some
extensions in second-order logic. Description logics are notable because most
of them are known to be decidable (but not always in a practically tractable
manner). In description logic, axioms are build from \emph{concepts},
\emph{roles} (that are unary or binary predicates) and \emph{individuals}. The
W3C OWL-DL standard is a widely-used language to describe domains with the
description logic.

Because description logics have been originally created from the perspective of
a \emph{knowledge representation language} and not a logic language, their
terminology (\emph{concept} or \emph{class}, \emph{role} or \emph{property},
\emph{individual},\ldots) is well-suited to knowledge description.

\emph{Modal logic}, that allows for statement qualification like
\emph{possibility} or \emph{necessity}, have been shown to be closely related
to description logics~\cite{Baader2001}. Modal logic allows to represent conveniently parallel
possible worlds and facts like ``the robot knows \emph{that the human knows}
how to read a recipe''.

\fxwarning{On modal logics, see the remark of McCarthy, in \cite{McCarthy2007}, section 3}

\emph{Temporal logic} are designed to represent and manipulate assertions whose
truth value may vary in time. We introduce one of its key idea (the
\emph{fluents}) later in the typology.

One last class of logics that is of particular relevance for robotic
applications is the \emph{probabilistic logics} or \emph{Bayesian logics}.
These logics provide a formal framework to reason on propositions whose truth
or falsity is uncertain. We elaborate below on the representation of uncertainty.

Note that most of these logic formalisms are still active research field on
their own, and practical considerations (especially the availability of
reasoners efficient enough for on-line use on a robot) often constrain the
choice of a logical formalism and a level of expressive power.

\subsubsection{Expressive power}

Logical formalisms each bring a certain level of expressive power. For
instance, the following classical syllogism can not be represented in
propositional logic because of the use of \emph{universal quantification}:

\begin{quote}
\begin{enumerate}
    \item All men are mortal,
    \item Socrates is a man,
    \item Therefore, Socrates is mortal
\end{enumerate}
\end{quote}

However, the following weak version of the syllogism can be represented in
propositional logic:

\begin{quote}
\begin{enumerate}
    \item If Socrates is a man, then Socrates is mortal,
    \item Socrates is a man,
    \item Therefore, Socrates is mortal
\end{enumerate}
\end{quote}

Generally speaking, expressive power comes at the cost of more complex
\emph{satisfiability} and \emph{consistency}\footnote{We precise these concepts
at section~\ref{sect|reasoning}.} computations, possibly leading to
untractable, if not undecidable (\ie systems where it is proven that a
proposition can not be decided to be true or false) problems.

\begin{figure}
    \centering
    \includegraphics[width=0.3\columnwidth]{expressive_overlap_dl_horn.pdf}
    \caption{expressiveness overlap of Description Logics and logic programs
    based on Horn clauses, taken from~\cite{Grosof2003}}
    \label{fig|overlap_dl_horn}
\end{figure}

Figure~\ref{fig|overlap_dl_horn} shows that the expressive power of description
logics and Horn clauses partially overlaps. In section~\ref{sect|reasoning} we
mention extensions to description logics based on rule systems that bring closer
the two approaches.

The relationships between expressive power and reasoning complexity that follow
has been extensively studied for Description Logics.
Zolin~\cite{ZolinDLComplexityNavigator} maintains a ``complexity navigator''
that allows to conveniently explore these relationships and indexes most of the
literature on that subject.

It can be noted that the relation between expressiveness and reasoning
complexity is fragile: for instance, adding the following axiom \stmt{farFrom
disjointProperty near} that states that two individuals can not be at the same
time near and far from each other, while looking semantically seemingly
innocuous, actually introduces what is called a \emph{complex role} that leads
to a much harder reasoning problem: concept satisfiability, for instance,
switches from a {\it NExpTime-complete} problem to a {\it NExpTime-hard}
problem (\ie, \emph{at least} as hard as the hardest problem in {\it
NExpTime}).

This kind of behaviours have practical consequences on run-time performances on
the robot, and require careful design of the knowledge structure to be mitigated.

\subsubsection{Open world and close world assumptions}
\label{sect|owa-cwa}

The \emph{close world} (CWA) vs. \emph{open world} (OWA) assumptions name a
modelling choice on the \emph{completeness} of a knowledge domain. In the close
world assumption, a proposition that can not be proven true is assumed to be
false (\emph{negation by failure}), while in the open world assumption, a
proposition may be considered either true, false or unknown.

This distinction is important in robotics where the robot may have to manipulate
concepts with only partial knowledge on them. For instance, let us imagine a robot
that sees a bottle on a table, whose bottom is hidden by another object. The
robot can not prove that the bottle is indeed \emph{on} the table. A knowledge
representation system relying on the closed world assumption would then assume
the bottle is \emph{not} on the table ($\lnot R^{CWA}_{isOn}(bottle, table)$)
whereas with the open world assumption, the proposition $R^{OWA}_{isOn}(bottle,
table)$ would be undecided. Example in table~\ref{table|cwa-owa-example} provides
another example of consequences of the CWA/OWA choice on reasoning.

\begin{table}
    \begin{center}
    \begin{tabular}{ll}
    {\bf Action} & {\bf Part involved} \\
    \hline
    {\tt PickSoftly} & hand \\
    {\tt PickAndPlace} & arm, hand \\
    {\tt MoveArm} & arm \\
    \hline
    \end{tabular}
    \end{center}
    \caption{Assuming the question is: \emph{select actions that do not require
    to move the arm}, a CWA reasoner would return {\tt PickSoftly} whereas an
    OWA reasoner would not return anything if the {\tt PickSoftly} action is
    not explicitly said not to involve the arm.}
    \label{table|cwa-owa-example}
\end{table}

The OWL language is specifically known to assume an open world.  Domains
constrained with the closed world assumption lead to more tractable inference
problems, and allow for instance the use of logic languages like Prolog. Thus,
several approaches exists to \emph{locally close} a domain (\cf
Levesque~\cite{Levesque2008}, section 24.3.2 for a summary of those ones).

\subsubsection{Representation of uncertainty and likelihood}

Sources of uncertainty for a robot are two-fold: uncertainty \emph{intrinsic}
to facts (like \emph{``It may rain tomorrow''}), uncertainty caused by
imperfect perception of the world (\emph{``Is the bottle really on the
table?''}). Most logics do not account explicitly for uncertainty. It must be
either relied on specific logics (like Bayesian logics) or on extensions of
classical logics.

\fxfatal{Representation of uncertainty: To be completed!}

\subsubsection{Meta-cognition: knowledge on the knowledge}

As stated by Cox and Raja~\cite{Cox2007}, meta-cognition is composed of both
\emph{``meta-level control of cognitive activities and the introspective
monitoring of such activities to evaluate and to explain them"}.

Sloman proposes in~\cite{Sloman2011} a detailed analysis of meta-cognition and
its different aspects in both natural (human) and artificial systems.

A knowledge representation system endowed with \emph{meta-cognition} is not
only able to manipulate knowledge but also to exhibit and manipulate the
structure of its knowledge and the reasoning process. For instance, the ability
to explain a logical inconsistency in a KRS is a meta-cognitive function, as is
the ability to expose and alter the knowledge structure (these two reasoning
techniques have their own entries in the taxonomy, at
section~\ref{sect|reasoning}.

At section~\ref{sect|introspection} below, we discuss the idea of
introspection.  Meta-cognition can be viewed as the technical facet of the
introspection in general.

@@

We do not discuss in this section the general strategies to construct a
knowledge model (they will be presented in
section~\ref{sect|design-strategies}). We focus here on questions that involve
representational challenges (time, space, context) or require specific
cognitive capabilities (theory of mind, introspection, memory).

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf B. Representation}
            child {node [taxon] {{\bf B.5}. Memory}}
            child {node [taxon] {{\bf B.4}. Introspection}}
            child {node [taxon] {{\bf B.3}. Modality}}
            child {node [taxon] {{\bf B.2}. Context}}
            child {node [taxon] {{\bf B.1}. Roles}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

\subsubsection{Role representations}

This section discusses strategies and approaches to represent in a knowledge
model three important \emph{roles}: spatial relations, time and representation
of actions.

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf B.1 Roles}
            child {node [taxon] {{\bf B.1.3}. Actions}}
            child {node [taxon] {{\bf B.1.2}. Time}}
            child {node [taxon] {{\bf B.1.1}. Space}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

\paragraph{Representation of space}

Symbolic representation of space is a widely studied topic. In particular, a
large literature corpus is available on spatial ontologies\fxwarning{Which one?}.

Two main classes of spatial relations are usually represented at the symbolic
level: the topology of environments and the placement of physical entities.

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf B.1.1 Space}
            child {node [taxon] {{\bf B.1.1.2}. Placement}}
            child {node [taxon] {{\bf B.1.1.1}. Topology}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

Topological maps are abstractions of an environment as graphs where nodes
represent places and edges represent connections between places. Because they
are symbolic representation, topological maps allow higher-level reasoning
(such as containment, connectivity, regions) than metric maps.

One important contribution to the building of a coherent representational stack
for space representation is the Spatial Semantic Hierarchy~\cite{Kuipers2000}
introduced by Kuipers. It consists in multiple interacting representations of
space, both qualitative and quantitative, that span from sensor-level
representation to ontologies of places and regions.

Symbolic representations of entities placement are either topological or
directional. The topological relation \concept{isOn}, for example, leads to
absolute statements: the validity of the relation is independent of the nature
of its subjects and objects, and is also independent of the observer. These
relations are also sometimes qualified as \emph{allocentric}.

On the contrary, the directional relation \concept{nextTo} is relative, and
depends on the relative size of the subject and the object. Two houses distant
of 2 meters from each other can be considered as next to each other. Two ants
separated by 2 meters are not next to each other.

The relation \concept{leftOf} is another example of relative spatial relation,
this time because it depends on the observer viewpoint. These relations are
also called \emph{egocentric}.

\fxwarning{\concept{front}/\concept{back} -> entities have or not a concept of front/back.}

Choices must be made in the knowledge representation system to adequately
represent relative spatial relations. Options include the computation of such
relations only on-demand, when the context is known (which viewpoints, etc.) or
storage of these facts in different models, one per agent.

\fxwarning{Find references on that subject -> classification of spatial
relations. Wikipedia?}

\paragraph{Representation of time}
\label{sect|time-representation}

As an agent acting at human-like time scale and dealing with temporal concepts
(like actions), a robot needs to represent and reason about
time. Time representation is split into two distinct abilities: representing
time points (both in the past -- which is roughly equivalent to assignment of
timestamps to events the robot perceives -- and in the future), and
representing passing time (situations, durations, timespans) like in
\emph{``the eggs will be cooked in 10 min''}.

This is usually formalised as a disjunction in time representation between
\emph{events}, that is, any durationless temporal concept, and
\emph{situations}, that is, any temporal concept with a non-zero
duration\footnote{Other definitions of a situation do exist, notably in the
context of \emph{situation calculus}~\cite{Levesque1998}, Reiter and Levesque
consider a situation to be an history of actions.}.

Numerous techniques to represent and reason about time have been devised.
Amongst the most significant ones, we can mention Allen's interval
algebra~\cite{Allen1984} and Ghallab's chronicles~\cite{Ghallab1996}.
\fxwarning{We could detail more chronicles and allen algebra...}
The concept of \emph{fluent} also play an important role for time
representation: fluents are properties (or \emph{conditions}) that change over
time, like $sees(agent1, apple, t)$.\fxwarning{A bit short... citation?}

We call a system that does not account for time (\ie that mentally permanently
lives in present) \emph{atemporal}.

\paragraph{Actions}

Events and actions are two temporal concepts that are of particular importance to
robotic systems, as systems that perceive, react and perform in their environment.

While representation of events boils down to label a timestamp, representation
of actions are more complex since not only they deal with durations, but they
also imply semantic interactions between concepts. From a taxonomy point of
view, actions are a particular type of event that normally leads to a situation
corresponding to the action realisation.

\emph{Thematic roles}~\cite{Gruber1965} (also found as \emph{semantic roles} or
\emph{theta roles} in the literature) allow to semantically qualify the
parameters of an action. The recipient of the action, the performer, the object
acted upon, the destination are some example of common thematic roles.
Table~\ref{table|theta-roles} presents a more comprehensive list of thematic
roles proposed by \cite{Aarts1997} (see~\cite{Gutierrez2001} for a comparison
of other sets of thematic roles present in the literature).

\begin{table}
\begin{center}

\begin{tabular}{lp{5cm}}

\toprule
       Role & Meaning \\
\midrule
      Agent & The \emph{doer} or instigator of the action denoted by the predicate. \\
    Patient & The \emph{undergoer} of the action or event denoted by the predicate. \\
      Theme & The entity that is moved by the action or event denoted by the predicate. \\
Experiencer & The living entity that experiences the action or event denoted by the predicate. \\
       Goal & The location or entity in the direction of which something moves. \\
Benefactive & The entity that benefits from the action or event denoted by the predicate. \\
     Source & The location or entity from which something moves. \\
 Instrument & The medium by which the action or event denoted by the predicate is carried out. \\
   Locative & The specification of the place where the action or event denoted by the predicate in situated. \\
\bottomrule

\caption{A list of thematic roles, as proposed by Aarts~\cite{Aarts1997}.
Depending on the action, only certain roles are meaningful.}

\label{table|theta-roles}
\end{table}

In the context of robotic, \emph{task} often represents an aggregate of
(atomic) actions. We use the term here as the representation of the abstract
model of an action. Tasks usually specify at least the conditions required to
performed the action and the consequences of the realization of the action, and
are central to the decisional layers of the robot, in particular for the
planning, monitoring and execution control activities. Knowledge representation
systems may thus have specific mechanisms to represent them (and possible, to
reason about them, as presented in section~\ref{sect|planning}).

Plans representation is closely related to action and task representation. It
is specifically discussed at section~\ref{sect|planning}.

\subsubsection{Context modeling}

Knowledge is contextualized information: it is essential for the robot to
associate the facts it represents to a \emph{context}. The context carries the
keys for the interpretation of the information and set a common ground for
interaction (and, in particular, communication~\cite{Jakobson1960}). It
implicitely defines the domain of validity of the facts and carries the
\emph{common-sense} knowledge required to fill the gaps of the knowledge
explicitely shared between the agents.

Context is generally difficult to recognise, represent and reuse because it is
multiform and largely implicit nature. It is also never unique: at a given
moment, several contexts, at different temporal, spatial, social scales,
overlap.

In the current literature in robotics and cognitive architectures, the term
\emph{context} usually simply refers to a set of beliefs that initiate a
representation (and reasoning) frame: in~\cite{Lemaignan2011a}, the robot
creates a context of interaction with a specific human by storing in a separate
model the beliefs of this human and using this knowledge when dialoging with
the human, the reasoning network is reinitialised in the GLAIR architecture
when the hypotheses that defined the current situation are not believed
anymore~\cite{Shapiro2009}.

This acception of \emph{context} is simplistic, and omit the overlapping and
multi-scale aspects of context modeling.

We see context representation as one of the main challenge of knowledge
representatio in general, and we will further discuss the importance and issues
brought by context modeling in the conclusion of the paper.

\subsubsection{Modality, contingency and theory of mind}
\label{sect|possible-worlds}

Linked to the context representation, but seen from another angle, knowledge
representation systems may support logical \emph{modality}. A knowledge model
is logically modal if is support the concept of \emph{possible worlds}, \ie,
parallel beliefs models (or \emph{interpretations}) that can be independently
accessed.

A \emph{contingent} proposition is defined as neither always true (a tautology)
nor always false (a contradiction) in every possible world: its truth value
depends on the context. In knowledge representation systems for robotics that
support logical modality, interpretations are often initialised with a common
set of initial beliefs (like common-sense knowledge). This initial common
knowledge is hence true in every possible worlds for the robot, and thus does
not belong to its contingent knowledge.

On the contrary, alternative mental models with contingent knowledge may be
used to represent different (possibly hypothetical or even imaginary) views on
the world, from the robot own perspective or context, or from other
perspectives computed by the robot.

The representation of the mental perspective of other agents has a particular
importance in human-robot interaction. It relies first on the ability to
literally \emph{view} the world from a standpoint which is not
\emph{egocentric}. This cognitive ability is referred as \emph{perspective
taking}. Flavell~\cite{Flavell1992} and Tversky~\cite{Tversky1999} define the
psychological grounds of perspective taking, that are themself originated in
Piaget's work on cognitive development (recent studies on infants include
Moll~\cite{Moll2006} for instance). Perspective taking begins to be studied on
robots as well~\cite{Trafton2005, Breazeal2006, Ros2010}.

The idea of a \emph{theory of mind}~\cite{Premack1978} emerges from the
perspective taking ability. It can be defined as the ability for one to
understand and acknowledge that other intelligent agents can have their own
mental state (that includes beliefs, intents, desires, knowledge) that is
possibly different from one's own. The \emph{attention} plays a central to the
development and recognition of a theory of mind~\cite{Baron-Cohen1985,
Leslie2000}.

A notable consequence of having a theory of mind is the representation of
\emph{false beliefs}, \ie, facts that are believed to be true for an agent, but
false for other ones. The \emph{Sally and Ann} experiment~\cite{Leslie2000}
(figure~\ref{fig|sally-ann}) is a classical example of a false belief
situation. In~\ref{fig|sally-ann-final}, Sally thinks the ball is in the beige
box because she did not see Ann moving it. An external observer asked ``Where
will Sally look for the ball?'' would answer ``in the blue box'' without a
theory of mind (\ie, a model of the knowledge of Sally), whereas it would
correctly answer ``In the beige box'' with a theory of mind.

\begin{figure}
        \centering
        \subfigure[Sally puts the ball in the beige box.]{
            \includegraphics[scale=0.3]{sally_ann1.pdf}
        } \hspace{8pt} %
        \subfigure[She goes away.]{
            \includegraphics[scale=0.3]{sally_ann2.pdf}
        } \\
        \subfigure[Ann moves the ball.]{
            \includegraphics[scale=0.3]{sally_ann3.pdf}
        } \hspace{8pt} %
        \subfigure[Where will Sally look for the ball?]{
            \includegraphics[scale=0.3]{sally_ann4.pdf}
            \label{fig|sally-ann-final}
        }
        \caption{The standard ``Sally and Ann'' false-beliefs experiment, taken
        from~\cite{Leslie2000}.}

        \label{fig|sally-ann}
\end{figure}

Scassellati~\cite{Scassellati2002} is one of the first to have implemented a
theory of mind on a humanoid robot.

\subsubsection{Self-knowledge: Who am I? What can I do?}
\label{sect|introspection}

\paragraph{Self-knowledge}

\emph{Self-knowledge} is the term used in philosophy and psychology to describe
the knowledge that an individual has, acquires or infers about itself through
its experiences. It answers the question ``What am I like?''.

Introspection is the technical meaning to access to self-knowledge by the
ability to self describe: what are my capabilities, what is my state
(performing some action, idling, etc.), what are my beliefs, what are my
intentions and my plans?

Introspection must be distinguished from meta-cognition: While introspection
may require meta-cognition (for instance to be able to expose its internal
knowledge), it is not always mandatory. The current state of the robot can be
represented as a simple instantiation of a specific category (for instance, if
the robot gives an object to the human, this state could be represented with the
triples \setstmt{robot performs action1,action1 isA Give}.

\paragraph{Modeling of the robot capabilities}

A particularly important aspect of self-knowledge for robots relates to the
description of its own capabilities: which sensors/actuators/computation
services exist and are currently available ?  While at a first level, these
descriptions can be static (\eg the robot has one laser scanner and two arms),
at more advanced levels, the description is updated and reflect the current
(and possibly past and future) state of the robot. Note that these descriptions
may also involve geometric descriptions (a kinematic chain, the pose of a
device, etc.) that may be deported outside of the main knowledge base. Efforts
trying to formalize, maintain and expose the capabilities and state of a robot
are not new (and ground themselves in work and techniques for self-descriptive
remote procedure calls in computing science), but take a renewed importance
with applications for high-level multi-robot cooperation.

Recent works by Kunze et al.~\cite{Kunze2011} seeks at defining a formal
language to represent the capabilities of a robot.

\subsubsection{Memory}
\label{sect|memory}

Memory has been studied at length in the cognitive psychology and
neuropsychology communities: Atkinson and Shiffrin~\cite{Atkinson1968}
introduce the idea of \emph{short-term} and \emph{long-term} memory,
Anderson~\cite{Anderson1976} splits memory into \emph{declarative} (explicit)
and \emph{procedural} (implicit) memories, Tulving~\cite{Tulving1985} organises
the concepts of \emph{procedural}, \emph{semantic} and \emph{episodic} memories
into a hierarchy. Short-term memory is refined with the concept of
\emph{working memory} by Baddeley~\cite{Baddeley2010}
(Figure~\ref{fig|memory_models}).

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{memory_models.pdf}
    \caption{Overview of the main types of memories, based
    on~\cite{Atkinson1968, Anderson1976, Tulving1985, Baddeley2010}}

    \label{fig|memory_models}
\end{figure}

It is worth emphasising that if memory is commonly associated to the process of
forgetting facts after a variable amount of \emph{time}, it actually covers
more mechanisms that are relevant to robotics, like selective remembering
triggered by a specific context or reinforcement learning.

Most knowledge representation systems offers some kind of memory as a pool of
facts that are not forgotten by the robot until it is halted (this memory is
often referred as a \emph{working memory}, but with a meaning unrelated to
Baddeley's definition). Some systems may propose persistent storages that allow
the robot knowledge to grow over time, while others may offer a larger range of
memory categories, like short term memory (that lasts for a couple of seconds)
or episodic memory (that allows the robot to selectively remember facts
associated to specific events).

In the larger field of cognitive architectures, the {\sc Soar}
architecture~\cite{Lehman2006} is one of those that tries to reproduce a
human-like memory organisation. The GLAIR cognitive architecture also have a
concept of long term/short term and episodic/semantic memories.

@@

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf C. Reasoning}
            child {node [taxon] {{\bf C.10}. Learning}}
            child {node [taxon] {{\bf C.9}. Naive physics}}
            child {node [taxon] {{\bf C.8}. Planning}}
            child {node [taxon] {{\bf C.7}. Prediction and explanation}}
            child {node [taxon] {{\bf C.6}. Presupposition}}
            child {node [taxon] {{\bf C.5}. Non-monotonicity}}
            child {node [taxon] {{\bf C.4}. Uncertainty}}
            child {node [taxon] {{\bf C.3}. Lazy evaluation}}
            child {node [taxon] {{\bf C.2}. Instantiation and structural alteration}}
            child {node [taxon] {{\bf C.1}. Standard reasoning}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

\subsubsection{Standard reasoning techniques}
\label{sect|std-reasoning}

We call \emph{standard reasoning techniques} techniques based on logical
inference, using resolution algorithms like \emph{forward chaining},
\emph{backward chaining} or \emph{semantic tableaux}.

Main reasoning problems include \emph{concept satisfiability},
\emph{consistency checking} and \emph{instance checking}.

Concept satisfiability verifies if it is possible to find a non-empty
\emph{interpretation} of a concept (or an expression defining a concept) in the
knowledge model. For instance, the formula \concept{Plant} $\land$
\concept{isRed}, which defines the concept of red plants, is satisfiable in a
model \concept{KB} iff $\exists a, $ \concept{Plant}$(a) \land$
\concept{isRed}$(a)$, \ie if we can find at least one red plant $a$ in our
model.

Checking the consistency of a model is equivalent to checking the
satisfiability of each of the concept defined in the knowledge model.

Instance checking consists in verifying that an individual $a$ is an
interpretation of a concept (or concept expression) $C$ in the knowledge model.
A typical example would be that we are provided with an instance
\concept{object1} and we want to know if this object is a kind of
\concept{Bottle} or \concept{Glass}.

Inferences can also be drawn from other constructs, whose availability depends
on the representation language. OWL, for instance, has constructs for:

\begin{itemize}
    \item class subsumption (to represent inheritance relations)

    \item reasoning on roles properties, including:
        \begin{itemize}
        \item entailments based on roles domain and range (for instance, if the
        domain of the role \concept{thinksOf} is known to be
        \concept{ThinkingAgent}, then \par \concept{thinksOf}$(a, b) \to
        $\concept{ThinkingAgent}$(a)$),

        \item universal, existential and cardinality constraints,

        \item several second-order predicates (inverse, symmetry, transitivity, etc.)

>>>>

\newcommand{\taxon}[2]{%
    \textbf{[#1] \emph{#2}}
}

@@

The paper is organized in three main sections.

Section~\ref{sect|features} introduces a typology of requirements and features
of knowledge representation systems for robotics, that we use as foundations to
compare and evaluate the systems we examine in this survey.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Evaluation Criteria: A Typology of Knowledge Representation Requirements for Robotics}
\label{sect|features}

We propose to organize requirements from this imaginary cooking scenario where
numerous desirable features for a knowledge representation system for robotics
have been identified without any particular order.

@@

As previously mentioned, we propose six main categories, that cover both basic
and practical aspects: the formal
\taxon{A}{Expressiveness} of the system, the breadth and nature
of the \taxon{B}{Representations} allowed by the system, the
\taxon{C}{Reasoning} capabilities, how knowledge it acquired
\taxon{D}{Acquisition}, how the system integrates with the other components of
the robot sotware architecture \taxon{E}{Integration} and finally what we call
the knowledge \taxon{F}{Instantiation}, \ie the actual content of the knowledge
base.

The next paragraph details only briefly the exact meaning of these categories
and introduce accordingly sub-categories. We kindly refer the interested reader
to~\cite{blabla} for in-depth discussion of each of these dimensions of
knowledge representation systems, as well as related literature.

Table~\ref{table|contribution-by-systems}, at the end of the article, summarizes
the main contributions of each of the surveyed KRS with respect to these
dimensions.

@@

@@

We subdivide the category \taxon{A}{Expressiveness} into five subcategories:
\taxon{A.1}{Logic formalism} classifies the underlying logical formalism of the
KRS (Horn clauses like for Prolog-based systems, Description
Logics~\cite{Baader2008}, Bayesian logics, etc.). \taxon{A.2}{Expressive Power}
quantifies the design choices made with respect to the actual expressive power
of the logical formalism: tradeoffs have to be found to allow both for high
expressiveness and tractable and decidable representations, and KRS authors
often add constraints on their representation logics to this end.
\taxon{A.3}{OWA/CWA} reflects the design choice between the \emph{open world}
(OWA) versus the \emph{closed world} (CWA) assumptions, which related to the
possibility of representing \emph{unknown} facts. \taxon{A.4}{Uncertainty}
relates to the ability for the system to represent both the \emph{intrinsic}
uncertainty of given facts, and the \emph{extrinsic} uncertainty due to
imperfect perception by the robot. \taxon{A.4}{} relates only to the possibility
of \emph{representation}. Reasoning on uncertain facts is discussed in
\taxon{C.4}{}. The last features of a KRS related to representation is
\taxon{A.5}{Meta-cognition} that conveys the capability for a system to
represent and exhibit it own internal structures and representation mechanisms.

@@

This category focuses on questions that involve
representational challenges (time, space, context) or require specific
cognitive capabilities (theory of mind, introspection, memory).
Note that the general strategies to actually \emph{construct} a
knowledge model are treated in section~\ref{sect|design-strategies}.

The dimension \taxon{B.1}{Roles}, with its sub-categories \taxon{B.1.1}{Space},
\taxon{B.1.2}{Time} and \taxon{B.1.3}{Actions}, deals with the approaches that
the KRS suggests to represent these three important \emph{roles} in robotics
(spatial relations, time and representation of actions). To give a few examples
(in no way exhaustive), \emph{Spatial relations} covers techniques like
topological maps, allocentric and egocentric frame projections. \emph{Time
representation} includes Allen's interval algebra~\cite{Allen1984}, Ghallab's
chronicles~\cite{Ghallab1996} or the concept of \emph{fluent}. \emph{Actions}
concern themselves with representation of \emph{pre-, post-conditions},
\emph{thematic roles} or \emph{plans}.

\taxon{B.2}{Context} describes if and how a given KRS represents the
\emph{interpretation frame} of the situation and its own knowledge. This is
often related to the initial \emph{common-sense knowledge} available to the
robot. \taxon{B.3}{Modality} refers to the ability, for the KRS, to represent
possible alternate worlds or belief states. A typical application of a
\emph{modal} knowledge model is the implementation of a \emph{theory of mind},
\ie an independent belief model for the other agents the robot interact with.
\taxon{B.4}{Introspection} covers both \emph{self-knowledge} (what do I know
about myself?) and a weaker version focused on the explicit representation and
reasoning on the \emph{robot's own skills} (can I execute a given task?).
Finally, \taxon{B.5}{Memory} refers to the implementation of
different kind of bio-inspired memory models like \emph{declarative},
\emph{procedural} or \emph{episodic}, and the attached processes
(\emph{forgetting}, \emph{selectively remembering}, etc.).

@@

As one can expect, many reasoning techniques born in artificial intelligence
have found their place in robotics over the years. We propose to consider ten
subcategories of \taxon{C}{Reasoning} that reflects in our opinion important
features in regards to the general requirements of robotics (like decision
making being tractable in pseudo-real time, being able to deal with partially
unknown or uncertain domain or specific reasoning related to interaction with
the physical world).

Again, the descriptions of these ten subcategories hereafter are very
superficial. Readers can refer to~\cite{blabla} for the details.

\taxon{C.1}{Standard reasoning} covers conventional reasoning techniques based
on logical inferences (for instance, \emph{forward/backward chaining} or
\emph{semantic tableaux}). The main problems that these techniques address are
\emph{concept satisfiability}, \emph{consistency checking} and \emph{instance
checking}. \taxon{C.2}{Instantiation and structural alteration} describes how
dynamic a system is with respect to its internal knowledge structure (allowing
for modification of the knowledge structure, often called \emph{TBox}, may
impact the capability for the system to \emph{learn} for instance).
\taxon{C.3}{Lazy evaluation} describes the ability for a KRS to delay active
knowledge acquisition or reasoning operations until the value is actually
needed, and may have a strong impact on the overall scalability of the system.
\taxon{C.4}{Uncertainty} groups all the features related to reasoning over
uncertain knowledge (like probabilistic inference).
\taxon{C.5}{Non-monotonicity} discusses how systems deal with non-monotonic
inferences, \ie exceptions to a rule or inferences that lead to the retractation
of previously inferred facts. Techniques to deal with non-monotonic reasoning
include for instance \emph{Default Logics}.  \taxon{C.6}{Presupposition
accomodation} is the ability for a system to automatically create a context
allowing to make sense of a proposition. This is tighly linked to the capability
for a system to deal with \emph{under-specification}. \taxon{C.7}{Prediction and
explanation} (with subcategories \taxon{C.7.1}{Projection},
\taxon{C.7.2}{Legality}, \taxon{C.7.3}{Diagnosis} and
\taxon{C.7.4}{Explanation}) relates to the capabilites for a system to predict
and project the outcomes of a sequence of actions onto the logical system, and
conversely, to justify and explain the sequence of entailments that lead to a
conclusion. Related to this category, \taxon{C.8}{Planning} discusses to which
extend a KRS can be used for \emph{symbolic task planning}, \ie the ability for
a robot to select a sequence of actions in order to reach a given final state.
\taxon{C.9}{Physics-based reasoning} describes the capability for a symbolic
knowledge representation system to rely on physical reasoning (which may include
geometry, dynamics, collision checking, etc.) to derive symbolic inferences.
Finally, \taxon{C.10}{Learning} groups the various capabilities that lead to
learning behaviours.
